---
title: "Homework 1"
author: "Md Kamrul Hasan Khan"
date: "February 17, 2017"
output: pdf_document
---
\begin{center}
\center{ \textbf{\textit{\underline{Answer to the Question Number 1}}}}
\end{center}
```{r}

(x<- c(9:16))

tail(x,3)

x[(x%%2 == 0)]

x[-which(x%%2 == 0)]

```

\begin{center}
\center{ \textbf{\textit{\underline{Answer to the Question Number 2}}}}
\end{center}
```{r}
n = 25
# for loop

sum = 0
for (i in 1:n)
{
  sum = sum + (0.5^i)
}
sum


# while loop

sum = 0
i <- 1
while (i <= n)
{
  sum = sum + (0.5^i)
  i = i+1
}
sum

# without using a loop
```

Sum of the Geometric series: 
$$\sum_{i=1}^{n}a^{i} = a\Big(\dfrac{1-r^n}{1-r}\Big)$$


```{r}
n = 25
a = 1/2
r = 1/2
sum <- a*((1-r^n)/(1-r))
sum
```
As n increases the sum tends to 1.

If n is large, without using loop program will be more robust to errors.

\begin{center}
\center{ \textbf{\textit{\underline{Answer to the Question Number 3}}}}
\end{center}
\textbf{Question I:} $x^3$ is $O(x^3)$ and $\Theta(x^3)$ but not $\Theta(x^4)$ 

\textbf{Answer:} $$\lim_{x\to\infty} \frac{x^3}{x^3} = 1>0 $$
So, $x^3$ is $O(x^3)$ and $\Theta(x^3)$.
$$\lim_{n\to\infty} \frac{x^3}{x^4} = \lim_{n\to\infty} \frac{1}{x} = 0$$
So, $x^3$ is not $\Theta(x^4)$.

\textbf{Question II:} For any real constants $a$ and $b>0$, we have 
$(n+a)^b = \Theta(n^b)$

\textbf{Answer:} $$\lim_{n\to\infty} \frac{(n+a)^b}{n^b} = \lim_{n\to\infty} \Big(\frac{n+a}{n}\Big)^b = \lim_{n\to\infty} \Big(1+\frac{a}{n}\Big)^b = 1>0 $$
So, $(n+a)^b = \Theta(n^b)$.

\textbf{Question III:} $(log(n))^k = O(n)$ for any k

\textbf{Answer:} $$\lim_{n\to\infty} \frac{(log(n))^k}{n} = 0$$
So, $(log(n))^k = o(n) = O(n)$ 

\textbf{Question IV:} $\frac{n}{n+1} = 1 + O(\frac{1}{n})$

\textbf{Answer:} $\frac{n}{n+1} = 1 + O(\frac{1}{n})$

$\Rightarrow \frac{n}{n+1} -1 =  O(\frac{1}{n})$

$\Rightarrow \frac{-1}{n+1} = O(\frac{1}{n})$

Now $\frac{-1}{n+1} = O(\frac{1}{n+1})$, thus $\frac{-1}{n+1} = O(\frac{1}{n})$

\textbf{Question V:} $\sum_{i=0}^{\lceil{log_{2} (n)\rceil}}2^i$ is $\Theta(n)$

\textbf{Answer:} Let $log_{2} (n) = k$

Now$$\sum_{i=0}^{k}2^i = \frac{2^{k+1}-1}{2-1} = 2*2^k - 1=2*2^{log_{2} (n)}-1=2n-1$$ 

So, $$\lim_{n\to\infty} \frac{2n-1}{n} =\lim_{n\to\infty}(2- \frac{1}{n})=2>0 $$

Therefore, $\sum_{i=0}^{\lceil{log_{2} (n)\rceil}}2^i$ is $\Theta(n)$


\vspace{.3in}

\begin{center}
\center{ \textbf{\textit{\underline{Answer to the Question Number 4(a)}}}}
\end{center}
```{r}
selsort <- function(A)
{
  n = length (A)
  for (i in 1:(n-1))
  {
    index = i
    for (j in (i+1):n)
    {
      if (A[j] < A[index])
        index = j
    }
    temp = A[i]
    A[i] = A[index]
    A[index] = temp
  }
  return (A)
}

x = sample(1:100, 10, replace = TRUE)
x
a <- selsort(x)
a
```
\begin{center}
\center{ \textbf{\textit{\underline{Answer to the Question Number 4(b)}}}}
\end{center}
```{r}
n = 100
y = rnorm (n)

# Mergesort code
ptm <- proc.time ()
mergearrays <- function(x,y){
  m = length(x)
  n = length(y)
  if(m==0){
    return(z = y)
    }
  if(n==0){
    return(z = x)
  }
  if (x[1]<=y[1]){
    return(z = c(x[1],mergearrays(x[-1],y)))
  }else{
    return(z = c(y[1],mergearrays(x,y[-1])))
  }
}
mergesort <- function(x){
  n = length(x)
  mid = floor(n/2)
  if(n > 1){
    return(mergearrays(mergesort(x[1:mid]),mergesort(x[(mid+1):n])))
  }else{
    return(x)
  }
}
merge<-mergesort(y)
time_mergesort = proc.time () - ptm
summary(time_mergesort)

# Bubble sort code
ptm <- proc.time ()
bubblesort <- function(A)
{
  n = length (A)
  repeat 
  { 
    swapped = FALSE
    for (i in 1:(n-1))
    {
      if (A[i] > A[i+1])
      {
        temp = A[i]
        A[i] = A[i+1]
        A[i+1] = temp
        swapped = TRUE
      }
    }
    if (swapped == FALSE) 
      break
  }
  return (A)
}
bubble<- bubblesort(y)
time_bubblesort = proc.time () - ptm
summary(time_bubblesort)
```
\begin{center}
\center{ \textbf{\textit{\underline{Answer to the Question Number 4(c)}}}}
\end{center}
```{r}
time_mergesort = rep (0 ,500)
time_bubblesort = rep (0 ,500)
for ( i in 1:500)
{
  ptm <- proc.time ()
  mergesort ( y )
  t1 = proc.time () - ptm
  time_mergesort [i] = t1 [["elapsed"]]
  ptm <- proc.time ()
  bubblesort ( y )
  t2 = proc.time () - ptm
  time_bubblesort [i]= t2 [["elapsed"]]
}
summary(time_mergesort)
summary(time_bubblesort)


```
Time complexities of mergesort and bubblesort are $O(nlog(n))$ and $O(n^2)$ respectively. So, for large n mergesort will take much much less time than bubblesort, which agrees the summary of the results.



