---
title: "Homework 2"
author: "Md Kamrul Hasan Khan"
date: "March 3, 2017"
output: pdf_document
---
\begin{center}
\textbf{\textit{\underline{Answer to the Question Number 1}}}
\end{center}
Given, $p(x) = sin(x) \hspace{1 in}$ [where $x \in (0,\frac{\pi}{2})$]

Now the cdf of $p(x)$ will be $P(x) = \int_{0}^{x} sin(x)dx= -cos(x)\Big|_0^x =1-cos(x)$

$\hspace{.45 in} \Rightarrow u = 1- cos(x)$ $\hspace{1 in}$ [where $u \sim Uinf(0,1)$]

$\hspace{.45 in} \Rightarrow cos(x) = 1-u$ 

$\hspace{.45 in} \Rightarrow x = arccos(1-u)$ 

```{r}
set.seed(7)
n = 10000
u = runif(n)
x = acos(u)
hist(x, freq = F, breaks = 30, col = "navajowhite4")
curve(sin(x), 0, pi/2, col = "blue", add = T, lw = 2)
```
```{r}
set.seed(7)
arsample <- function(dist, minsup, maxsup, maxdist, n)
{
  rv = rep(0,n)
  for (i in 1:n)
  {
    while(1)
    {
      x = runif(1, minsup, maxsup)
      y = runif(1, 0, maxdist)
      if (y <= dist(x))
      {
        rv[i] = x
        break
      }
    }
  }
  return(rv)
}

sinfun <- function(x)
{
  return(sin(x))
}

x = arsample(sinfun, 0, pi/2, 1, 10000)
hist (x, freq = F, breaks = 30, col ="navajowhite4")
curve (sinfun, add = T, col = "maroon1")
```
\begin{center}
\textbf{\textit{\underline{Answer to the Question Number 2}}}
\end{center}
Given, $p(x) = \lambda x^{-\lambda-1}$ where $x \in [1,\infty)$ and $\lambda \geq 2$

So, the cdf will be

$P(x) = \int_{1}^{x}\lambda x^{-\lambda-1} dx = -x^{-\lambda}\Big|_1^x= 1-x^{-\lambda}$

Let, $u = 1-x^{-\lambda}$ $\hspace{1 in}$ [where $u \sim Uinf(0,1)$]

$\Rightarrow x^{-\lambda} = 1-u$

$\Rightarrow x = (1-u)^{-1/\lambda}$

```{r}
set.seed(7)
pareto <- function(lambda)
{
  u = runif(n)
  x = 1/u^(1/lambda)
}
n = 100
x = pareto(3)
hist(x, breaks = 30, freq = F, col = "navajowhite4", xlim = c(1,5))

fy <- function(y,lambda) lambda*(y^(-lambda-1))
curve(fy(x,3), add = T, col = "blue", lw = 2)

```
Mean

$E(X^{2.736})= \int_{1}^{\infty} x^{2.736}\lambda x^{-\lambda-1} dx = \lambda \int_{1}^{\infty} x^{-\lambda+1.736} dx = \lambda \frac{x^{-\lambda+2.736}}{-\lambda+2.736}\Bigg|_1^\infty= 3 \frac{x^{-3+2.736}}{-3+2.736}\Bigg|_1^\infty=3 \frac{1}{-0.264x^{0.264}}\Bigg|_1^\infty = - 0 +\frac{3}{0.264}= 11.36$

```{r}
set.seed(12345)
lambda = 3
mean(pareto(lambda)^2.736)
```





